Towards a Machine Learning approach for Aggregate Computing

### 2021-11-06 paper-2021-woa-ac-ml.tex
add figure of aggregate program learning loop, hop count and the overall learning program

\section{Aggregate Computing with Reinforcement}
In this section, we describe our first result in Aggregate Computing and Machine learning combination.
% 
Hence we briefly describe how Reinforcement Learning can be injected into the computation loop,
 and what kind of Reinforcement Learning can be used with Aggregate Computing.

\subsection{Reinforcement Learning integration}
From a local view, program execution can be seen as:
$$
f : \textit{Context} \rightarrow \textit{Output}
$$
Where the context contains:
\begin{itemize}
  \item $\sigma(name) \rightarrow \textit{data}$: a function from sensor name to the perceived value
  \item $\phi(\textit{ID}) \rightarrow \textit{Export}$: a function from id to neighbourhoods export data
\end{itemize}
The computational field $\theta$ contains all output produced by the system.

@@

For our first tests, 
 we decide to set up an Independent learning technique, by which each agent try to improve our current local policy.
 Our approach is partially inspired by Swarm Q-Learning~\cite{nguyen2018swarm} and leverage Hysteretic Q-Learning~\cite{hysteretic-q} to 
 handle
 the non-stationarity in the environment.
%
In particular, each agent maintains a local Q table ($Q_i(s, a)$, where i is the agent id) following the standard Hysteretic update:
$$
\delta \leftarrow r + \gamma * max_a Q_i(s', a) - Q_i(s_t, a_t)
$$
$$
Q(s_t, a_t) =  \begin{cases} 
  Q(s_t, a_t) + \alpha * \delta & \mbox{if } delta >= 0 \\ 
  Q(s_t, a_t) + \beta * \delta & \mbox{otherwise }
\end{cases}
$$

At the end of the episodes, each agent has collected its local experience, which can vary in different zones. So, we combine the N q table
 selecting the maximum value for each state-action pair:
$$
Q_{max}(s, a) = max_i(Q_i(s, a))
$$
Then, in the new episode, each agent starts with the $Q_max$ and try to refine it. 
 This max combination works because we assume a homogeneous behaviour, where each agent perform
 in the same way when reaching the same state.

\section{Evaluation}
In this section, we outline the results of the Reinforcement Learning application in one simple
 aggregate program: the hop count.

@@

 \begin{minted}{scala}
def space(hopCount: Int, k: Int, w: Int) {
  val currentSpeed = speed(hopCount, k)
  recentValues(currentSpeed, w)
}
\end{minted}

Where \textbf{\texttt{k}} and \textbf{\texttt{w}} are two parameters of this algorithm.
%
The policy is learnt locally, during simulation, using the Q-Learning algorithm. 
 The reward function consists in: 
\[
\systeme*{0 : \textit{hop count output is right}, -1 : \textit{otherwise}}
\]
This is a common reward function used to maintain the nodes in a stable situation or to
 fasten the time needed to reach a stable one.
%
So, the learning loop, can be defined as:
\begin{minted}[obeytabs=true,tabsize=1]{scala}
val inf = Double.PositiveInfinity
rep((initState, initPolicy, inf)) {
  (oldState, policy, hopCount) => 
      val currentState = 
        state(hopCount, k, w)
      val action = policy(state)
      val output = mux(source) 
        { 0 } 
        { minHood(hopCount) 
          + 1 
          + action
        } 
      }
      val currentReward = 
        reward(output)
      val improvedPolicy = 
        qLearning(oldState, 
          state, 
          action, 
          reward,
          policy
      (currentState, policy, output)
    )
}
\end{minted}

>>>>

\section{Aggregate Computing with Reinforcement}
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{img/algorithm-learning.pdf}
  \caption{This figure shows the main steps of reinforcement learning in aggregate computing.
  Each agent starts with his copy of a q-table. 
  Then, during the simulation, they refine them q-tables changing q-values according to their experience (different colours mean different values). 
  In the end, a new $Q_{max}$ table will be created aggregating each agent table. 
  Then, a new episode starts using this $Q_{max}$ as a starting point.
  }
  \label{fig:aggregate-q-learning}
\end{figure}
In this section, we describe our first result in Aggregate Computing and Machine learning combination.
% 
Hence we briefly describe how Reinforcement Learning can be injected into the computation loop,
 and what kind of Reinforcement Learning can be used with Aggregate Computing.

From a local view, program execution can be seen as:
$$
f : \textit{Context} \rightarrow \textit{Output}
$$
Where the context contains:
\begin{itemize}
  \item $\sigma(name) \rightarrow \textit{data}$: a function from sensor name to the perceived value
  \item $\phi(\textit{ID}) \rightarrow \textit{Export}$: a function from id to neighbourhoods export data
\end{itemize}
The computational field $\theta$ contains all output produced by the system.

@@

For our first tests, 
 we decide to set up an Independent learning technique, by which each agent try to improve our current local policy (shown in \Cref{fig:aggregate-q-learning}).
 Our approach is partially inspired by Swarm Q-Learning~\cite{nguyen2018swarm} and leverage Hysteretic Q-Learning~\cite{hysteretic-q} to 
 handle
 the non-stationarity in the environment.
%
In particular, each agent maintains a local Q table ($Q_i(s, a)$, where i is the agent id) following the standard Hysteretic update:
$$
\delta \leftarrow r + \gamma * max_a Q_i(s', a) - Q_i(s_t, a_t)
$$
$$
Q(s_t, a_t) =  \begin{cases} 
  Q(s_t, a_t) + \alpha * \delta & \mbox{if } delta >= 0 \\ 
  Q(s_t, a_t) + \beta * \delta & \mbox{otherwise }
\end{cases}
$$

At the end of the episodes, each agent has collected its local experience, which can vary in different zones. So, we combine the N q table
 selecting the maximum value for each state-action pair:
$$
Q_{max}(s, a) = max_i(Q_i(s, a))
$$
Then, in the next episode, each agent starts with the $Q_{max}$ and try to refine it.
 This max combination works because we assume a homogeneous behaviour, where each agent perform in the same way when reaching the same state.

\section{Evaluation}
\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{img/hop-count.png}
  \caption{Graphical result of hop count evaluation. The red node is the target zone. The label 
  associated with each node is the result of the aggregate program evaluated in that zone.}
  \label{fig:hop-count}
\end{figure}
In this section, we outline the results of the Reinforcement Learning application in one simple
 aggregate program: the hop count (\Cref{fig:aggregate-q-learning}).

@@

 \begin{minted}{scala}
def state(hopCount: Int, k: Int, w: Int) {
  val currentSpeed = speed(hopCount, k)
  recentValues(currentSpeed, w)
}
\end{minted}

Where \textbf{\texttt{k}} and \textbf{\texttt{w}} are two parameters of this algorithm.
%
The policy is learnt locally, during simulation, using the Q-Learning algorithm. 
 The reward function consists in: 
\[
\systeme*{0 : \textit{hop count output is right}, -1 : \textit{otherwise}}
\]
This is a common reward function used to maintain the nodes in a stable situation or to
 fasten the time needed to reach a stable one.
%
The \Cref{fig:aggregate-learning-loop} shows the learning loop in details.
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{img/aggregate-program-evaluation.pdf}
  \caption{todo}
  \label{fig:aggregate-learning-loop}
\end{figure}
%\begin{minted}[obeytabs=true,tabsize=1]{scala}
%val inf = Double.PositiveInfinity
%rep((initState, initPolicy, inf)) {
%  (oldState, policy, hopCount) => 
%      val currentState = 
%        state(hopCount, k, w)
%      val action = policy(state)
%      val output = mux(source) 
%        { 0 } 
%        { minHood(hopCount) 
%          + 1 
%          + action
%        } 
%      }
%      val currentReward = 
%        reward(output)
%      val improvedPolicy = 
%        qLearning(oldState, 
%          state, 
%          action, 
%          reward,
%          policy
%      (currentState, policy, output)
%    )
%}
%\end{minted}
